---
title: "R Notebook"
output: html_notebook
---


```{r Read Data and Clean it}

library(tidyverse)
library(lubridate)

#here we read and clean data by filling "?" with NA values,
#converting columns to dates and times, as well as converting
#character columns to numeric values and then filtering to 
#the correct date range 
data <-read.table("household_power_consumption.txt", sep = ";",
                  header = TRUE) %>% 
       as_tibble %>% 
       na_if("?") %>%
       mutate(across(.cols = -(1:2),
                     .fns = as.numeric)) %>% 
       mutate(across(.cols = 1, .fns = dmy)) %>% 
       mutate(across(.cols = 2, .fns = hms)) %>% 
       filter(Date >= ymd("2007-02-01") & Date <= ymd("2007-02-02"))


#need to create a datetime column with mutate
data <- data %>% mutate(dt_tm = make_datetime(year(Date),
                                              month(Date),
                                              day(Date),
                                              hour(Time),
                                              minute(Time),
                                             second(Time)))

#need to customize and set tickmarks for the plot. We see the plot has
#thurs, fri, and sat. Which correspond to Feb 1, Feb 2, Feb 3. at time 00:00:00
#we will make it a vector and pass it to the breaks arg of scale_x_continuous,
#then call labels to set the day of week
xticks <- c(make_datetime(2007,2,1,0,0,0),
            make_datetime(2007,2,2,0,0,0),
            make_datetime(2007,2,3,0,0,0))
```
```{r Make Plot 1}

p1 <- ggplot(data = data)+ 
        geom_histogram(mapping = aes(x = Global_active_power), fill = "red",
                       col = "black", bins = 24) +
        labs(x = "Global Active Power (kilowatts)", y = "frequency",
             title = "Global Active Power") + 
        theme_classic() +
        ylim(c(0,1250))

ggsave("Plot1.png", dpi = 1200)
```
```{r Plot 2}


p2 <- ggplot(data = data) + 
        geom_line(mapping = aes(x = dt_tm, y=Global_active_power),
                  color = "black") +
        labs(x = "Day of week", y = "Global Active Power",
             title = "Daily Fluctuations") +
        scale_x_continuous(breaks = xticks,labels = c("Thurs", "Fri", "Sat")) +
        theme_bw()

ggsave("plot2.png", dpi = 1200)

```

```{r Plot 3}



#sub meter is not tidy form. can pivot to get it in tidy form and
#is much better for ggplot
piv <- data %>% pivot_longer(cols = 7:9, names_to = "sub_meter_type",
                              values_to = "key")


 p3 <- ggplot(data = piv) +
          geom_line(mapping = aes(x = dt_tm, y = key,color = sub_meter_type)) +
          labs(x = "Day of week", y = "Energy sub metering") +
          scale_x_continuous(breaks = xticks, labels = c("Thurs", "Fri", "Sat"))

ggsave("plot3.png", dpi = 1200)

```

```{r Plot 4 multiplot}

p4 <- ggplot(data = data) + geom_line(mapping = aes(x = dt_tm, y = Voltage)) +
                            scale_x_continuous(breaks = xticks, labels = 
                                                    c("Thurs", "Fri", "Sat")) +
                            labs(x = "Day of week")

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}


multiplot(p1,p2,p3,p4, cols = 2)
ggsave(plot = multiplot(p1,p2,p3,p4, cols = 2),"plot4.png", dpi = 1200)
```








Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
